Comparison of approaches to using machine learning models to increase the efficiency of the global search algorithm for solving multicriterial problems

### 2024-08-09 2024/OPTIMA/OPTIMA_2024_MCO.tex
fix all the comments except two

@@

In each series of experiments for these optimization algorithms, the coefficients $\alpha$ from (\ref{eq:16}) changed in the range $[0.1, 0.9]$. Comparison with the algorithm MGSA that does not use machine learning methods (i.e. with $\alpha = 0$ in the characteristic (\ref{eq:16})) and with the algorithm ML\_MGSA\_Dist taking into account the distance to the separating hyperplane was carried out. The comparison was made on the base of  the HV index characterizing quality of the obtained Pareto set approximation \cite{ML_MCO_2023,Evtushenko2014}. Further subsections include the results of only those calculations in which the $\alpha$ parameter provided the best quality of the solution found. Tables contain results of solving all the problems (\ref{eq:07}) with fixed accuracy $\varepsilon=0.01$ in the termination condition (\ref{eq:14}). Figures present results for a series of decreasing $\varepsilon$ which generates increasing numbers of trials.

@@

Acceleration of the algorithm (in terms of reducing the number of search trials required to evaluate the Pareto set) is achieved through the application of machine learning methods. Namely, when using the SVM method, trial points in the criteria space are classified (each point is/is not an efficient solution), on the basis of which a numerical estimate of the probability of belonging to the Pareto set is built. Polynomial and radial basis function kernels in SVM allows one to adequately approximate Pareto front of a complex structure.

>>>>

\textcolor[rgb]{0,0,1}{\textit{Remark 1.} The Lipschitz constant of the objective function is assumed a priori unknown; it is adaptively estimated during the search.
The value $M$ from (\ref{eq:10}) can be used in Lipschitz constant estimation.\\
\textit{Remark 2.} It follows from relationship between the Lipschitz and H{\" o}lder constants of the functions $\varphi(y)$ and $\varphi(y(x))$, as well as from the theorem of convergence of the global search algorithm (see \cite{Strongin2000}), that parameter $r$ from (\ref{eq:10}) should satisfy the necessary condition $r>2^{2-1/N}$. 
In the test experiments described in Section~\ref{sec:5} we used the value $r=3$, which ensured the correct solution of the entire series of problems. In applied optimization problems the value $r=3$ also ensures the solution in the most cases.
}

@@

In each series of experiments for these optimization algorithms, the coefficients $\alpha$ from (\ref{eq:16}) changed in the range $[0.1, 0.9]$. Comparison with the algorithm MGSA that does not use machine learning methods (i.e. with $\alpha = 0$ in the characteristic (\ref{eq:16})) and with the algorithm ML\_MGSA\_Dist taking into account the distance to the separating hyperplane was carried out. The comparison was made on the base of the HV index characterizing quality of the obtained Pareto set approximation. 
\textcolor[rgb]{0,0,1}{The HyperVolume (HV) index corresponds to the volume of the region dominated by the Pareto set approximation and bounded by a given reference point \cite{ML_MCO_2023,Evtushenko2014}. The larger the hypervolume index, the better is the corresponding approximation.
}

Further subsections include the results of only those calculations in which the $\alpha$ parameter provided the best quality of the solution found. Tables contain results of solving all the problems (\ref{eq:07}) with fixed accuracy $\varepsilon=0.01$ in the termination condition (\ref{eq:14}). Figures present results for a series of decreasing $\varepsilon$ which generates increasing numbers of trials.

@@

Acceleration of the algorithm (in terms of reducing the number of search trials required to evaluate \textcolor[rgb]{0,0,1}{the Pareto set approximation with a certain precision}) is achieved through the application of machine learning methods. Namely, when using the SVM method, trial points in the criteria space are classified (each point is/is not an efficient solution), on the basis of which a numerical estimate of the probability of belonging to the Pareto set is built. Polynomial and radial basis function kernels in SVM allows one to adequately approximate Pareto front of a complex structure.
